---
title: "Querying dbGaPdb"
author: "David McGaughey"
date: "8/16/2017"
output: html_document
---
```{r results='hide',message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE)
```

# Load libraries
```{r}
library(dbGaPdb)
library(RSQLite)
library(dplyr)
library(dbplyr)
library(stringr)
```
# Downloading the dbGaPdb sqlite file and connecting to it via RSQLite
We strongly suggest you provide a path for the `pull_dbGaPdb_sqlite()` script so you know exactly where the sqlite file is. The below is taken from the quick start from the readme
```{r}
dbGaPdb_file = 'dbGaPdb.sqlite'
if(!file.exists(dbGaPdb_file))
  dbGaPdb_file <- pull_dbGaPdb_sqlite(local_path='~/')
dbGaPdb <- src_sqlite(dbGaPdb_file)
# If you've already downloaded the sqlite file
## dbGaPdb <- src_sqlite('~/PATH/your_dbGaP.sqlite')
```

# Bits
The dbGaPdb metadata is stored as a sqlite file. It is updated each week, so if you want the latest metadata, then re-run `pull_dbGaPdb_sqlite()`. Otherwise you can use your downloaded sqlite file indefinitely. As of 2017-08-16, the sqlite file is ~108 mb compressed and ~547 mb uncompressed. It must be uncompressed to use.

# What tables are available?
```{r}
dbGaPdb
```

## dbplyr
We will be using dplyr via dbplyr to query the sql database. This allows for, in our opinion, a gentler introduction to querying sqlite for people who don't have previous experience writing sql queries. A good tutorial on dplyr and sql is located [here](http://www.datacarpentry.org/R-ecology-lesson/05-r-and-databases.html)

## SQL experts
Use this convention to run sql queries directly on the sqlite database
```{r}
tbl(dbGaPdb, sql('SELECT study_accession, dataset_accession, short_description FROM study_dataset_info LIMIT 5' ))
```

# Link to two of the tables
```{r}
study_info_SQL <- tbl(dbGaPdb, "study_info")
study_variable_info_SQL <- tbl(dbGaPdb, "study_variable_info")
```

# Peek
```{r}
study_info_SQL %>% head()
study_variable_info_SQL %>% head()
```

# Filter to find number of variables with more than 10,000 females
```{r}
study_variable_info_SQL %>% 
  filter(as.numeric(female_count) > 10000)
```

# Filter and group by studies to find number of studies where there are any variables with > 10,000 females
```{r}
study_variable_info_SQL %>% 
  filter(as.numeric(female_count) > 10000) %>% 
  group_by(study_accession) %>% 
  summarise(`Number of variables with >10,0000 females`=n())
```

# Aggregate by version-less studies
You may have noticed that there are many versions (e.g phs000200.v4 and phs000200.v5). If you want to aggregate or study at the study level, we'll need to remove the .vNUM part with a mutate. To do this operation we have to convert the data to a data frame to run the rowwise() and mutate() function. But now we can see that we have `study_variable_info_SQL %>% filter(as.numeric(female_count) > 10000) %>% data.frame() %>% rowwise() %>% mutate(study_accession2 = str_split(study_accession, '\\.')[[1]][1]) %>% group_by(study_accession2) %>% summarise(`Number of variables with >10,0000 females`=n()) %>% nrow()` unique studies. 
```{r}
study_variable_info_SQL %>% 
  filter(as.numeric(female_count) > 10000) %>% 
  data.frame() %>% 
  rowwise() %>% 
  mutate(study_accession2 = str_split(study_accession, '\\.')[[1]][1]) %>% 
  group_by(study_accession2) %>% 
  summarise(`Number of variables with >10,0000 females`=n())
```

# What diseases are studied across the studies?
```{r}
tbl(dbGaPdb, "study_info") %>% 
  select(root_study_accession, disease) %>% 
  data.frame() %>% 
  rowwise() %>% 
  mutate(study_accession2 = str_split(root_study_accession, '\\.')[[1]][1]) %>% 
  select(study_accession2, disease) %>% 
  unique()
```

# Loading whole tables into memory as a R data frame
You cannot run the full gamut of dplyr or base R queries on `tbl(SQLite)`, as several (e.g. tail(), sample_n()) require the data to be in a data frame (in memory). As the database isn't not too huge, most computers can handle having the tables loaded as a data frame. After this you have a data frame and can run any base R or dplyr queries
```{r}
# study_variable_info <- study_variable_info_SQL %>% data.frame()
study_variable_info_SQL %>% 
  select(study_accession, description) %>% 
  collect() %>% 
  sample_n(10)
```


