---
title: "Linking to NIH Common Data Elements"
author: "Robert Butler"
date: "August 16, 2017"
output: html_document
---

## Quick Start

Getting your json batch from CDE -> import to dbGaPdb (parse with python)-> list of studies (phs study accession) which have your data type:

```
code goes here
```

## NIH Common Data Elements (CDE)

NIH is currently producing Common Data Elements (cde.nih.nlm.gov) to harmonize the categories of data present across studies available in their databases. This is currently implemented in dbGaP, and provides both a common naming terminology (often an amorphous concept) and *consensus measures for Phenotypes and eXposures (PhenX)* identifier. You can download a cross reference to dbGaP an therefore SRA.

```{r cars}
summary(cars)
```

## Generating Data Table from exported json  
Currently there is not programatic access to CDE elements, so a batch file must be downloaded. You can go to the NIH CDE website https://cde.nlm.nih.gov/cde/search?selectedOrg=PhenX and search their PhenX terms for dbGaP. From the results you can stratify by category (on the left) and export a json (upper right corner) 

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

##Importing Table to dbGaPdb

texttexttext

##Joining with variable tables for more information

##Generating a list of Studies with your data type
