---
title: "Schema"
author: "Michelle Miron"
date: "8/16/2017"
output: html_document
---

# Schema

## What does dbGaP metadata contain?

#### It contains five files:
1. study_dataset_info.txt 
2. study_info.txt
3. study_variable_code_value.txt
4. study_variable_info.txt
5. study_dataset_info.txt

### Each file has different study-level info:
1. study_dataset_info.txt

|      |row_names | study_id| study_version|study_accession | dataset_id| dataset_version|dataset_accession |name                         |dataset_type_category |description                                                                                                                                                                |short_description                                                                                                                                                          |
|:-----|:---------|--------:|-------------:|:---------------|----------:|---------------:|:-----------------|:----------------------------|:---------------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
|15726 |15726     |      470|             4|phs000470.v4    |       2636|               2|pht002636.v2      |TARGET_RT_Subject_Phenotypes |subject               |The data tables contains Rhabdoid Tumor (RT) subject phenotypes including gender (n=1 variable). The complete dataset has not been submitted, so more data is forthcoming. |The data tables contains Rhabdoid Tumor (RT) subject phenotypes including gender (n=1 variable). The complete dataset has not been submitted, so more data is forthcoming. |
|4724  |4724      |        7|            27|phs000007.v27   |        303|               5|pht000303.v5      |icam3_1s                     |subject               |Intercellular Adhesion Molecule 1, Generation 3 Cohort Exam 1                                                                                                              |Intercellular Adhesion Molecule 1, Generation 3 Cohort Exam 1                                                                                                              |
|13483 |13483     |      285|             3|phs000285.v3    |       1564|               2|pht001564.v2      |A4F06BH                      |subject               |1985-1986, Year 0. Nutrient Data, Dietary History. Cardiovascular disease and its risk factors among young adult participants.                                             |1985-1986, Year 0. Nutrient Data, Dietary History. Cardiovascular disease and its risk factors among young adult participants.                                             |
|7582  |7582      |       80|            12|phs000080.v12   |        171|               2|pht000171.v2      |psel1_7s                     |subject               |P-Selectin, Offspring Cohort Exam 7                                                                                                                                        |P-Selectin, Offspring Cohort Exam 7                                                                                                                                        |
|4325  |4325      |        7|            26|phs000007.v26   |        194|               5|pht000194.v5      |brach3_1s                    |subject               |Non-invasive Vascular Brachial Test, Generation 3 Cohort Exam 1                                                                                                            |Non-invasive Vascular Brachial Test, Generation 3 Cohort Exam 1                                                                                                            |
2. study_info.txt

|     |row_names | root_study_id| root_study_version|root_study_accession_p |root_study_accession | this_study_id| this_study_version|this_study_accession_p |this_study_accession |name                                                                  |disease                      |
|:----|:---------|-------------:|------------------:|:----------------------|:--------------------|-------------:|------------------:|:----------------------|:--------------------|:---------------------------------------------------------------------|:----------------------------|
|2461 |2461      |           707|                  1|phs000707.v1.p1        |phs000707.v1         |           707|                  1|phs000707.v1.p1        |phs000707.v1         |Next Generation Mendelian Genetics: Hereditary Neurological Disorders |Nervous System Diseases      |
|1684 |1684      |             7|                 17|phs000007.v17.p6       |phs000007.v17        |           282|                  6|phs000282.v6.p6        |phs000007.v17        |NHLBI Framingham Candidate Gene Association Resource (CARe)           |Cardiovascular Diseases      |
|1418 |1418      |             7|                  9|phs000007.v9.p4        |phs000007.v9         |           194|                  4|phs000194.v4.p4        |phs000007.v9         |Framingham SHARe VEGF Study                                           |Cardiovascular System        |
|2031 |2031      |           218|                 12|phs000218.v12.p2       |phs000218.v12        |           463|                  9|phs000463.v9.p2        |phs000218.v12        |TARGET: Acute Lymphoblastic Leukemia (ALL) Pilot Phase 1              |Acute Lymphoblastic Leukemia |
|2042 |2042      |           218|                  9|phs000218.v9.p1        |phs000218.v9         |           464|                  6|phs000464.v6.p1        |phs000218.v9         |TARGET: Acute Lymphoblastic Leukemia (ALL) Expansion Phase 2          |Acute Lymphoblastic Leukemia |

3. study_variable_code_value.txt

|        |row_names |study_id | study_version|study_accession | variable_id| variable_version|variable_accession |code |value                                                |rank_order |
|:-------|:---------|:--------|-------------:|:---------------|-----------:|----------------:|:------------------|:----|:----------------------------------------------------|:----------|
|2865133 |2865133   |280      |             1|phs000280.v1    |       92915|                1|phv00092915.v1     |2    |Plaque only                                          |2          |
|2345598 |2345598   |134      |             6|phs000134.v6    |       55151|                1|phv00055151.v1     |NA   |UNKNOWN                                              |3          |
|55302   |55302     |7        |            13|phs000007.v13   |       55197|                1|phv00055197.v1     |NA   |UNKNOWN OR ARTIFACT OR TEST STOPPED                  |2          |
|2162768 |2162768   |21       |             2|phs000021.v2    |       35016|                1|phv00035016.v1     |0    |Absent (or insufficient evidence to rate as present) |1          |
|669001  |669001    |7        |            21|phs000007.v21   |        1004|                1|phv00001004.v1     |0    |NO                                                   |2          |
 
 4. study_variable_info.txt
 |        |row_names |study_id | study_version|study_accession | dataset_id| dataset_version|dataset_accession |dataset_name             | variable_id| variable_version|variable_accession |name        |units |reported_type |calculated_type |dataset_type_category |male_count |female_count |max  |min |sd     |non_null_count |null_count | valid_value_count|description                                                                         |
|:-------|:---------|:--------|-------------:|:---------------|----------:|---------------:|:-----------------|:------------------------|-----------:|----------------:|:------------------|:-----------|:-----|:-------------|:---------------|:---------------------|:----------|:------------|:----|:---|:------|:--------------|:----------|-----------------:|:-----------------------------------------------------------------------------------|
|430704  |430704    |7        |            29|phs000007.v29   |         19|               3|pht000019.v3      |ex0_17s                  |        3230|                1|phv00003230.v1     |FJ336       |NA    |NA            |enum_integer    |subject               |799        |1223         |1799 |31  |977.6  |2022           |7          |              2022|OTHER VASCULAR DIAGNOSIS                                                            |
|1129516 |1129516   |285      |             2|phs000285.v2    |       1594|               2|pht001594.v2      |AFLWUP1                  |      114076|                2|phv00114076.v2     |FY012JOB    |NA    |NA            |integer         |subject               |823        |938          |2    |1   |0.4639 |3206           |416        |              3206|FOUND JOB & STAY IN FOR 2 YRS - MON 12                                              |
|1230056 |1230056   |287      |             5|phs000287.v5    |       3699|               1|pht003699.v1      |SHHS1_PSG                |      197495|                1|phv00197495.v1     |asa15       |NA    |NA            |enum_integer    |subject               |512        |689          |618  |583 |24.75  |1201           |4330       |              1201|In last two weeks, did you take any aspirin or aspirin containing meds?             |
|349130  |349130    |7        |            27|phs000007.v27   |       4811|               1|pht004811.v1      |t_physactf_ex02_3b_0914s |      229316|                1|phv00229316.v1     |mvpmintue15 |NA    |NA            |decimal         |subject               |1371       |1607         |54   |NA  |3.707  |5956           |NA         |              5956|MVPA Minutes Tuesday Hour: 15 (KV944738728) ih: mvpmintue15                         |
|73840   |73840     |7        |            16|phs000007.v16   |        748|               2|pht000748.v2      |pat1_8s                  |       72634|                1|phv00072634.v1     |PDONE       |NA    |NA            |enum_integer    |subject               |1321       |1578         |2523 |88  |1352   |2899           |NA         |              2899|Was PAT protocol completed? (Determined at time of scan or at time of interpreting) |

5. study_dataset_info.txt
|      |row_names | study_id| study_version|study_accession | dataset_id| dataset_version|dataset_accession |name                               |dataset_type_category |description                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                            |short_description                                                                                                                                                                                                                                                                        |
|:-----|:---------|--------:|-------------:|:---------------|----------:|---------------:|:-----------------|:----------------------------------|:---------------------|:------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
|4984  |4984      |        7|            27|phs000007.v27   |       5164|               1|pht005164.v1      |t_ctvasclng_2011_m_0952s           |subject               |Vascular Morphology Measurements from CT Lung Image (Round 2), Offspring, Generation 3, Omni 1, and Omni 2, 2008-2011. This dataset contains vasculature measurements of the whole lung from the second round CT scans with the primary goal of accessing coronary artery calcification.                                                                                                                                                                                                                                                                                                                                                                                                               |Vascular Morphology Measurements from CT Lung Image (Round 2), Offspring, Generation 3, Omni 1, and Omni 2, 2008-2011. This dataset contains vasculature measurements of the whole lung from the second round CT scans with the primary goal of accessing coronary artery calcification. |
|1044  |1044      |        7|            16|phs000007.v16   |        390|               2|pht000390.v2      |eye0_1975s                         |subject               |Eye Examination, Original Cohort 1973 - 1975                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                           |Eye Examination, Original Cohort 1973 - 1975                                                                                                                                                                                                                                             |
|14726 |14726     |      342|             9|phs000342.v9    |       2946|               2|pht002946.v2      |Framingham_SHARe_Sample_Attributes |sample                |The sample attributes data table includes shareid, SHARe sample ID for 100K, 550K, and OMNI5.0 genotype samples, sample analyte type (DNA), body site where samples were collected, plate and well information, high flag, and DNA amplification status.                                                                                                                                                                                                                                                                                                                                                                                                                                               |The sample attributes data table includes shareid, SHARe sample ID for 100K, 550K, and OMNI5.0 genotype samples, sample analyte type (DNA), body site where samples were collected, plate and well information, high flag, and DNA amplification status.                                 |
|11241 |11241     |      209|            12|phs000209.v12   |       2197|               1|pht002197.v1      |MESA_MesaFamily_EyeRefraction      |subject               |MESA Family Eye Ancillary Study Dataset: This dataset provides eye refraction measures for MESA Family participants.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   |MESA Family Eye Ancillary Study Dataset: This dataset provides eye refraction measures for MESA Family participants.                                                                                                                                                                     |
|7663  |7663      |       86|             1|phs000086.v1    |        234|               1|pht000234.v1      |DCCT_m9excaif                      |subject               |The information published as &#39;Adverse Events and their Association with Treatment Regimens in the DCCT&#39; (for published results see: &lt;a href=&quot;http://www.ncbi.nlm.nih.gov/pubmed/8722064?ordinalpos=2&amp;itool=EntrezSystem2.PEntrez.Pubmed.Pubmed_ResultsPanel.Pubmed_DefaultReportPanel.Pubmed_RVDocSum&quot;&gt; PMID 8722064&lt;/a&gt;) is represented in nine datasets (prefixed with: m9-).  M9excaif is one of four datasets representing variables used to analyze the rate of specific adverse events occurrences in CSII pump versus MDI users (both groups are part of the experimental group).  M9excaif data focus on occurrences of catheter infections in those groups. |DCCT Adverse Events - Rate Comparison (Catheter Infection: CSII vs MDI)                                                                                                                                                                                                                  |



```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
